{\fontfamily{ptm}\selectfont Linear Convergence of Stochastic Nesterov's Accelerated Proximal Gradient method under Interpolation Hypothesis, Truth or Just a Dream?}

### 2025-07-17 Notebook LaTEX/SAGD Under Interpolation/main.tex
>w<, changes

            Suppose an inexact evaluation of proximal gradient operator at $x$ yield an approximation $\tilde x$, characterized by: 

@@

                & \exists: w \in \partial \left[

@@

                \right](\tilde x)\; \text{s.t: } \Vert w\Vert \le \epsilon \Vert x - \tilde x\Vert. 

@@

@@

            Suppose that, a random variable $\tilde \nabla f(x)$ is used to estimate the proximal gradient operator and, it produced an error $w$, and the estimate $\tilde x$ as defined in \ref{lemma:inex-pg-ineq-proto}. 
            Let $B \ge 0$ be constant such that $D_f(\tilde x,x)\le B/2 \Vert x - \tilde x\Vert^2$. 

@@

            Suppose that in addition, there exists $\epsilon \ge 0$ such that: 

@@

            Then it makes the following inequality true: 

@@

            In here, $z, x$ is not a random variable, $\epsilon$ just a constant, but it's determined by $z$ and $x$, one of the obvious choice for it in this lemma is

@@

            Of course, look, if there is no random variable and $\tilde \nabla f(x)$ is simply not an probailistic estimate then the expectation is gone and, it has: 

@@

        Suppose that $\tilde \nabla f(x)$ is an estimate of the true gradient $\nabla f(x)$ with relative error $\epsilon \ge 0$. 
        Then, the inexact proximal gradient operator is defined by: 

@@

            T_B^{(\epsilon)}(x | F) = \argmin_{z \in \RR^n}\left\lbrace

@@

        And it satisfies 
            \left\Vert \tilde \nabla f(x) - \nabla f(x)\right\Vert = \epsilon \Vert x - \tilde x\Vert. 

@@

    \begin{definition}[stochastic proximal gradient operator with relative error]

@@

        Suppose that $\tilde \nabla f(x)$ is a random variable which is an estimate of the true gradient $\nabla f(x)$ and, let the constant $\epsilon \ge 0$ be defined as: 

@@

            \epsilon &=

@@

            \right]}. 

@@

            \widetilde T_B^{(\epsilon)}(x| F) = \argmin_{z \in \RR^n}\left\lbrace

@@

@@

            & x_k =  T_{L}^{(\epsilon_k)}(y_k | F_{I_k}) \text{ or }, \widetilde T_{L}^{(\epsilon_k)}(y_k | F_{I_k})\\

@@

@@

@@

        Let $\expect_{k}$ be the expectation conditioned on all $\tilde \nabla f(y_i)$ for $i = 0, 1, \ldots, k - 1$. 
        We note that under this conditioning, the iterates $x_{k - 1}, v_{k - 1}, y_k$ are not random variables, but $x_k$, and $v_k$ are. 

@@

        We use Lemma \ref{lemma:stoch-pg-ineq} with $x = y_k, z = z_k, \tilde x = x_{k}$ and, $B = L$ then it has: 

@@

            F(z_k) - \expect_{k - 1} F(x_k) 

@@

            + \frac{\epsilon_k}{2}\expect_{k - 1} \left[\Vert y_k - x_k \Vert^2\right] 

@@

                - \frac{L - \epsilon_k}{2}\expect_{k - 1} \left[\Vert z_k - x_k\Vert^2\right]

@@

            \alpha_k F(\bar x) + (1 - \alpha_k)F(x_{k - 1}) - \expect_{k - 1} F(x_k) - \frac{\mu\alpha_k(1 - \alpha_k)}{2}\Vert \bar x - x_{k - 1}\Vert^2

@@

                + \frac{\epsilon_k}{2}\expect_{k - 1} \left[\Vert y_k - x_k \Vert^2\right] 
                - \frac{L - \epsilon_k}{2}\expect_{k - 1} \left[\Vert z_k - x_k\Vert^2\right]

@@

            \alpha_k F(\bar x) + (1 - \alpha_k)F(x_{k - 1}) - \expect_{k - 1} F(x_k) 

@@

                + \frac{\epsilon_k}{2}\expect_{k - 1} \left[\Vert y_k - x_k \Vert^2\right] 
                - \frac{L - \epsilon_k}{2}\expect_{k - 1} \left[\Vert z_k - x_k\Vert^2\right]
            \\&\underset{\textcolor{red}{(?)}}{\le}
            \alpha_k F(\bar x) + (1 - \alpha_k)F(x_{k - 1}) - \expect_{k - 1} F(x_k) 

@@

            (1 - \alpha_k)(F(x_{k - 1}) - F(\bar x)) + F(\bar x) - \expect_{k - 1}F(x_k)

@@

                + F(\bar x) - \expect_{k - 1}F(x_k) 
                - \frac{L - \epsilon_k}{2}\expect_{k - 1} \left[\Vert z_k - x_k\Vert^2\right]

@@

            &\underset{\circled{2}}{=} 

@@

                + F(\bar x) - \expect_{k - 1}F(x_k) 
                - \frac{\alpha_k^2(L - \epsilon_k)}{2}\expect_{k - 1} \left[\Vert v_k - \bar x\Vert^2\right]. 

@@

@@

            & \expect_{k - 1}F(x_k) - F(\bar x) - \frac{\alpha_k^2(L - \epsilon_k)}{2}\expect_{k - 1} \left[\Vert v_k - \bar x\Vert^2\right]

@@

            \right). 

@@

>>>>

            Suppose an inexact evaluation of proximal gradient operator at $x$ yield an approximation $\tilde x$ such that: 

@@

                & (\exists w)(\exists \epsilon ): w \in \partial \left[

@@

                \right](\tilde x), \Vert w\Vert \le \epsilon \Vert x - \tilde x\Vert. 

@@

            The smallest possible choice for $\epsilon$ when $x \neq \tilde x$ is $\epsilon = \Vert w\Vert/\Vert x - \tilde x\Vert$ and, if $x = \tilde x$ then $\epsilon = 0$ is the smallest. 

@@

            Suppose that, $\tilde \nabla f(x)$ is a random variable which estimates $\nabla f(x)$, cause proximal gradient operator to produce error $w$, and estimate $\tilde x$ as given by \ref{lemma:inex-pg-ineq-proto}. 
            Assume there exists $B \ge 0$ be constant such that $D_f(\tilde x,x)\le B/2 \Vert x - \tilde x\Vert^2$. 

@@

            If in addition, there exists an $\epsilon \ge 0$: 

@@

            Then it has: 

@@

            In here, $z, x$ is not a random variable, $\epsilon$ just a constant, but it's determined by $z$ and $x$. 
            Assuming $z \neq \tilde x$ and, $\tilde x \neq x$, then one of the smallest choice for it in this lemma is

@@

            Otherwise, $\epsilon = 0$ is an obvious choice and, it's the smallest. 
            It depends on both $z$ and $x$. 
            Of course, look, if there is no random variable and $\tilde \nabla f(x)$ is simply not a probabilistic estimate then the expectation is gone and $x\neq \tilde x$, it has: 

@@

        Suppose that $\tilde \nabla f(x)$ estimates $\nabla f(x)$. 
        We define the inexact proximal gradient operator by the relationships between: 
            \item $\tilde x = \mathbf T_B^{(\epsilon)}(x | F)$ is an inexact output of proximal gradient operator by evaluating on $\tilde \nabla f(x)$. 
            \item $B \ge 0$ is any constant such that it satisfies $D_f(\tilde x, x) \le B/2\Vert \tilde x - x\Vert^2$. 
            \item $\epsilon \ge 0$ is a constant that quantifies the error of inexact evaluation. 
        Then, we define the error $\epsilon$ by: 

@@

            \epsilon = \begin{cases}
                \frac{\left\Vert \tilde \nabla f(x) - \nabla f(x)\right\Vert}{\Vert x - \tilde x\Vert} & \text{if } x \neq \tilde x, 
                0 & \text{else } x = \tilde x. 
            \end{cases}
        And the inexact output is defined as: 
            \tilde x = \mathbf T_B(x | F) = \argmin_{z \in \RR^n}\left\lbrace

@@

@@

    \begin{definition}[stochastic proximal gradient operator with relative error]\label{def:stoch-pg-opt-rel-err}

@@

        Suppose that $\tilde \nabla f(x)$ is random variable and, it estimates $\nabla f(x)$. 
        Then stochastic proximal gradient operator are the relationships between
            \item $\tilde x= \mathbf{\widetilde T}_B(x| F)$, an inexact output of proximal gradient operator by evaluating on $\tilde \nabla f(x)$. 
            \item $\epsilon$ an error which is pre-determined constant. 
            \item Any $B \ge 0$ such that it satisfies $D_f(\tilde x, x) \le B/2\Vert x - \tilde x\Vert^2$. 
        Here, $\epsilon$ satisfies: 

@@

            \epsilon 
            \begin{cases}

@@

                \right]} & \text{if } x \neq \tilde x \;\wedge\; z \neq \tilde x
                0 & \text{else }
            \end{cases}

@@

            \tilde x = \mathbf{\widetilde T}_B(x| F) = \argmin_{z \in \RR^n}\left\lbrace

@@

        For this definition of the stochastic proximal gradient operator, Lemma \ref{lemma:stoch-pg-ineq} is applicable. 

@@

            & x_k =  \mathbf T_{L}(y_k | F) \text{ or }, \mathbf{\widetilde T}_L(y_k | F)\\

@@

    The following definition gives a momentum sequence where it makes the derivation of the convergence rate easier. 
    \begin{definition}[relaxed momentum sequence]\label{def:relax-momen-seq}
        Let $(\alpha_k)_{k \ge 0}$ be non-negative sequence. 
        Let $L, \mu$ be some constant such that $L > \mu \ge 0$. 
        It is a relaxed momentum sequence if the following conditions are satisfied: 
            \item $\alpha_0 \in (0, 1]$ and for all $k \ge 1$, it satisfies that $\alpha_k \in (\mu/L, 1)$. 
        In the context of its usage, the constants $L, \mu$ are the Lipschitz smoothness constant  and, strong convexity constant associated with a smooth and strongly convex function. 

    % ------------------------------------------------------------------------------------------------------------------
    \begin{lemma}[properties of the sequence]\label{lemma:seq-properties}
    \subsection{building up the convergence results}
        In this section we derive a generic convergence results. 

@@

            Suppose the iterates sequence $(y_k, x_k, v_k)_{k\ge 0}$ are generated by algorithms satisfying Definition \ref{def:SNAG}. 
            Assume that
                \item it uses stochastic proximal gradient operator as in Definition \ref{def:stoch-pg-opt-rel-err}, 
                \item it is initialized with $v_{-1} = x_{-1}$, $\alpha_0 = 1$ and, the momentum sequence $(\alpha_k)_{k \ge 0}$ satisfies Definition \ref{def:relax-momen-seq}. 
            Define $\expect_{k}$ to be the expectation conditioned on $\tilde \nabla f(y_i)$ for $i = 1, 2, \ldots, k - 1$. 
            Then for all $k \ge 1$, for all $\bar x \in \RR^n$ it satisfies the inequality: 
                & \expect_kF(x_k) - F(\bar x) - \frac{\alpha_k^2(L - \epsilon_k)}{2}\expect_k \left[\Vert v_k - \bar x\Vert^2\right]
                \right) + \frac{\epsilon_k}{2}\expect_k \left[\Vert y_k - x_k \Vert^2\right].  
            Define $z_k = \alpha_k \bar x + (1 - \alpha_k)x_{k - 1}$.
            Then, the error sequence $(\epsilon_k)_{k \ge 0}$ is given by: 
                (\forall k \ge 1)\; \epsilon_k = 
                \frac{\expect_k\left[
                        \tilde \nabla f(y_k) - \nabla f(y_k)
                    \right\Vert
                    \Vert \bar x - v_{k}\Vert
                    \expect_k\left[
                        \Vert \bar x - v_{k}\Vert
                        \Vert z_k - y_k\Vert
                }. 

@@

            Recall $\expect_{k}$ is the expectation conditioned on all $\tilde \nabla f(y_i)$ for $i = 0, 1, \ldots, k - 1$. 
            We note that under this conditioning the only random variable is $\tilde \nabla f(y_k)$, so iterates $x_{k - 1}, v_{k - 1}, y_k$ are not random variables, but $x_k$, and $v_k$ are. 

@@

            We use Lemma \ref{lemma:stoch-pg-ineq} with $x = y_k, z = z_k, \tilde x = x_{k}$ and, $B = L$ then it means: 

@@

                F(z_k) - \expect_k F(x_k) 

@@

                + \frac{\epsilon_k}{2}\expect_k \left[\Vert y_k - x_k \Vert^2\right] 

@@

                    - \frac{L - \epsilon_k}{2}\expect_k \left[\Vert z_k - x_k\Vert^2\right]

@@

                \alpha_k F(\bar x) + (1 - \alpha_k)F(x_{k - 1}) - \expect_k F(x_k) - \frac{\mu\alpha_k(1 - \alpha_k)}{2}\Vert \bar x - x_{k - 1}\Vert^2

@@

                    + \frac{\epsilon_k}{2}\expect_k \left[\Vert y_k - x_k \Vert^2\right] 
                    - \frac{L - \epsilon_k}{2}\expect_k \left[\Vert z_k - x_k\Vert^2\right]

@@

                \alpha_k F(\bar x) + (1 - \alpha_k)F(x_{k - 1}) - \expect_k F(x_k) 

@@

                    + \frac{\epsilon_k}{2}\expect_k \left[\Vert y_k - x_k \Vert^2\right] 
                    - \frac{L - \epsilon_k}{2}\expect_k \left[\Vert z_k - x_k\Vert^2\right]
                \\&\underset{\circled{2}}{\le}
                \alpha_k F(\bar x) + (1 - \alpha_k)F(x_{k - 1}) - \expect_k F(x_k) 

@@

                (1 - \alpha_k)(F(x_{k - 1}) - F(\bar x)) + F(\bar x) - \expect_kF(x_k)

@@

                    + F(\bar x) - \expect_kF(x_k) 
                    - \frac{L - \epsilon_k}{2}\expect_k \left[\Vert z_k - x_k\Vert^2\right]
                    + \frac{\epsilon_k}{2}\expect_k \left[\Vert y_k - x_k \Vert^2\right] 

@@

                &\underset{\circled{3}}{=} 

@@

                    + F(\bar x) - \expect_kF(x_k) 
                    - \frac{\alpha_k^2(L - \epsilon_k)}{2}\expect_k \left[\Vert v_k - \bar x\Vert^2\right]
                    + \frac{\epsilon_k}{2}\expect_k \left[\Vert y_k - x_k \Vert^2\right]. 

@@

            \textcolor{red}{Explanations for \circled{1}, \circled{2}, \circled{3} are not done yet}. 

@@

                & \expect_kF(x_k) - F(\bar x) - \frac{\alpha_k^2(L - \epsilon_k)}{2}\expect_k \left[\Vert v_k - \bar x\Vert^2\right]

@@

                \right) + \frac{\epsilon_k}{2}\expect_k \left[\Vert y_k - x_k \Vert^2\right]. 

@@

\subsection{Discussion}
